---
title: "Untitled"
author: "Bingquan Wu"
date: "December 7, 2018"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,
                      warning = FALSE)
```


```{r,fig.width=10,fig.height=7}
mydata <- read.csv(file="d:/columbia/EDAV/project/2015_Street_Tree_Census_-_Tree_Data.csv",
                   header=TRUE, sep=",")

library(ggplot2)
library(dplyr)
library(gridExtra)
library(grid)
library(lattice)
th<-theme(plot.title = element_text(size=22))+
  theme(axis.text=element_text(size=12),axis.title=element_text(size=16))

boroughrootstone<-mydata %>% group_by(borough) %>% count(root_stone) %>% 
  mutate(proportion = n / sum(n))
boroughrootstone2<- subset(boroughrootstone,boroughrootstone$root_stone == "Yes" ) 
boroughrootgrate<-mydata %>% group_by(borough) %>% count(root_grate) %>% 
  mutate(proportion = n / sum(n))
boroughrootgrate2<- subset(boroughrootgrate,boroughrootgrate$root_grate == "Yes" ) 
boroughtrunkwire<-mydata %>% group_by(borough) %>% count(trunk_wire) %>% 
  mutate(proportion = n / sum(n))
boroughtrunkwire2<- subset(boroughtrunkwire,boroughtrunkwire$trunk_wire == "Yes" ) 
boroughtrunklight<-mydata %>% group_by(borough) %>% count(trnk_light) %>% 
  mutate(proportion = n / sum(n))
boroughtrunklight2<- subset(boroughtrunklight,boroughtrunklight$trnk_light == "Yes" ) 
boroughbrchlight<-mydata %>% group_by(borough) %>% count(brch_light) %>% 
  mutate(proportion = n / sum(n))
boroughbrchlight2<- subset(boroughbrchlight,boroughbrchlight$brch_light == "Yes" ) 
boroughbrchshoe<-mydata %>% group_by(borough) %>% count(brch_shoe) %>% 
  mutate(proportion = n / sum(n))
boroughbrchshoe2<- subset(boroughbrchshoe,boroughbrchshoe$brch_shoe == "Yes" ) 

s1 <- ggplot(boroughrootstone2,aes(x=borough))+
  geom_bar(aes(y=proportion),stat="identity",fill="#56B4E9")+
  labs(title="root problem caused by stone",x="borough", y = "Proportion")
s2 <- ggplot(boroughrootgrate2,aes(x=borough))+
  geom_bar(aes(y=proportion),stat="identity",fill="#56B4E9")+
  labs(title="root problem caused by grate",x="borough", y = "Proportion")
s3 <- ggplot(boroughtrunkwire2,aes(x=borough))+
  geom_bar(aes(y=proportion),stat="identity",fill="#CC79A7")+
  labs(title="trunk problem caused by wire",x="borough", y = "Proportion")
s4 <- ggplot(boroughtrunklight2,aes(x=borough))+
  geom_bar(aes(y=proportion),stat="identity",fill="#CC79A7")+
  labs(title="trunk problem caused by light",x="borough", y = "Proportion")
s5 <- ggplot(boroughbrchlight2,aes(x=borough))+
  geom_bar(aes(y=proportion),stat="identity",fill="#009E73")+
  labs(title="branch problem caused by light",x="borough", y = "Proportion")
s6 <- ggplot(boroughbrchshoe2,aes(x=borough))+
  geom_bar(aes(y=proportion),stat="identity",fill="#009E73")+
  labs(title="branch problem caused by shoe",x="borough", y = "Proportion")
grid.arrange(s1, s2, s3, s4, s5, s6, ncol=2)
```

From the plot above, we can observe many patterns.\
We can see that among the 6 problems, root problems caused by stone has a much higher percentage than other problem, around 20% of trees has this problems, while almost no trees has branch problem caused by shoe. \
And we can see that Manhattan has a much higher percentage of perblem in root problem caused by grate and truck problem caused by light, while a much lower problem proportion in brach problem caused by light.\
Also, we can see that Staten Island has a relatively lower proportion on almost every problem meaning that it has a good preservation and management of the trees.\

