---
title: "Final Project"
author: "Xiaoyu Chen"
date: "12/9/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,
                      message = FALSE,
                      tidy.opts=list(width.cutoff=50),tidy=TRUE)
```

### 1. Tree numbers for each borough by year

```{r}
library(tidyverse)

mydata <- read_csv("/Users/xiaoyu/Desktop/Final/NYC_Trees.csv")
mydata_1995 <- read_csv("/Users/xiaoyu/Desktop/Final/1995.csv")
mydata_2005 <- read_csv("/Users/xiaoyu/Desktop/Final/2005.csv")

mydata <- mydata %>% mutate(health = replace_na(health, "Death"))
borough_2015 <- data.frame(table(mydata$borough))
colnames(borough_2015)[colnames(borough_2015) == 'Var1'] <- 'Borough'
borough_2015$Borough <- as.character(borough_2015$Borough)
borough_1995 <- data.frame(table(mydata_1995$Borough))
colnames(borough_1995)[colnames(borough_1995) == 'Var1'] <- 'Borough'
borough_1995$Borough <- as.character(borough_1995$Borough)
borough_2005 <- data.frame(table(mydata_2005$boroname))
colnames(borough_2005)[colnames(borough_2005) == 'Var1'] <- 'Borough'
borough_2005$Borough <- as.character(borough_2005$Borough)
borough_2005$Borough[borough_2005$Borough == "5"] <- "Staten Island"
borough_1995["Year"] <- "1995"
borough_2005["Year"] <- "2005"
borough_2015["Year"] <- "2015"

borough <- rbind(borough_1995, borough_2005, borough_2015)
colnames(borough)[colnames(borough) == 'Freq'] <- 'Numbers'
```

```{r}
borough %>% ggplot(aes(x = Year, y = Numbers, group = Borough, color = Borough)) + 
              geom_line() + ggtitle("Tree Numbers for Each Borough by Year") + 
                theme(plot.title = element_text(size=22)) + 
                  theme(axis.text=element_text(size=12),axis.title=element_text(size=16))
```

For five boroughs in New York City, number of trees varying by year intuitively reflect the green rate in each borough. It is evident to see from graph that number of trees show a trend of increasing. For the absolute tree numbers, *Queens* ranks first in five boroughs, following by *Brooklyn*, *Staten Island*, *Bronx* and *Manhatten*. While for the increasing rate, *Brooklyn* ranks first, following by *Bronx*, *Manhattan*, *Queens* and *Staten Island*.

### 2. Top five species numbers by year

```{r, results="hide"}
sort(table(mydata$spc_common), decreasing=TRUE)

name2015 <- c('London planetree', 'honeylocust', 'Callery pear', 'pin oak', 'Norway maple')
freq2015 <- c(87014, 64264, 58931, 53185, 34189)
spc2015 <- data.frame(name2015, freq2015)
colnames(spc2015)[colnames(spc2015) == 'name2015'] <- 'Species'
colnames(spc2015)[colnames(spc2015) == 'freq2015'] <- 'Numbers'
spc2015["Year"] <- "2015"

sort(table(mydata_1995$Spc_Common), decreasing=TRUE)
freq1995 <- c(88040, 33727, 31293, 36553, 109321)
spc1995 <- data.frame(name2015, freq1995)
colnames(spc1995)[colnames(spc1995) == 'name2015'] <- 'Species'
colnames(spc1995)[colnames(spc1995) == 'freq1995'] <- 'Numbers'
spc1995["Year"] <- "1995"

sort(table(mydata_2005$spc_common), decreasing=TRUE)
freq2005 <- c(89529, 52191, 63593, 43895, 74721)
spc2005 <- data.frame(name2015, freq2005)
colnames(spc2005)[colnames(spc2005) == 'name2015'] <- 'Species'
colnames(spc2005)[colnames(spc2005) == 'freq2005'] <- 'Numbers'
spc2005["Year"] <- "2005"

spc <- rbind(spc1995, spc2005, spc2015)
spc$Species <- as.character(spc$Species)
spc$Numbers <- as.integer(spc$Numbers)
```

```{r}
spc %>% ggplot(aes(x = Year, y = Numbers, group = Species, color = Species)) + 
          geom_line() + ggtitle("Top Five Species Numbers by Year") + 
            theme(plot.title = element_text(size=22)) + 
              theme(axis.text=element_text(size=12),axis.title=element_text(size=16))
```

In year 2015, the top five tree species in New York city is *London planetree*, *honeylocust*, *Callery pear*, *pin oak* and *Norway maple*. Each of five species presented a different growing trend in the past ten year. For example, the growing trend of *London planetree* is very stable. Knowing that the *Norway maple* is very invasive, the NYC government pulled seedlings of *Norway maple* from moist soil before they get too large to prevent them from destroying native ecosystems, causing trouble in yards and gardens, and creating visual blight. Therefore, the number of *Norway maple* decreased dramatically in the past ten years, varying from the first to the last.

### 3. Tree_dbh by health
```{r}
g <- ggplot(mydata, aes(tree_dbh)) + scale_fill_brewer(palette = "Spectral")
g + geom_histogram(aes(fill=health), 
                   binwidth = .7, 
                   col="black", 
                   size=.1) +  xlim(0,30) + ggtitle("Tree_dbh by Health") +
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(plot.title = element_text(size=22)) +
  theme(axis.text=element_text(size=12),axis.title=element_text(size=16))
```

The graph descirbes the health status of each diameter of the trees in the dataset. It is evident that most trees are in *good* status and obviously, dead trees account only for a small fraction. From the plot, we also know that most trees fall in diameters below 20 inches and the status for each tree is ranked by *Good*, *Fair*, *Poor* and *Death*.