---
title: "Hmk4-Q3"
author: "Bingquan Wu"
date: "November 6, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,
                      warning = FALSE)
```

## Diameter at breast height of tree from NYC -- tree_dbh

I choose the Diameter at breast height of tree -- tree_dbh as my variable.\
Data is by downloading the csv file from the following website\
https://data.cityofnewyork.us/Environment/2015-Street-Tree-Census-Tree-Data/uvpi-gqnh

```{r}
mydata <- read.csv(file="d:/columbia/EDAV/project/2015_Street_Tree_Census_-_Tree_Data.csv",
                   header=TRUE, sep=",")
tree_dbh<-mydata['tree_dbh']
library(ggplot2)
ggplot(mydata, aes(x=tree_dbh)) + geom_histogram(color="black", fill="white")
```

From the histogram we can see the distribution of the tree's diameter, and that most tree have a diameter from 0 to 30. And lots of them are in the range of 10-20.\
While from the scale of x-axis, we can now that there is few tree that has diameter of more than 400. And they determined the x-axis value.\
So we have a common question to ask, how much trees can be considered as outlier, from the histogram, we can know that tree has a diameter of more 50 is rare and can be considered to be outlier.

```{r}
ggplot(mydata, aes(y=tree_dbh)) + geom_boxplot()
```

By plotting a boxplot. We can clearly see the statistic of the distribution, and find out the outlier easily.we can see no trees with diameter negative, which is for sure, but we need to check since the histogram have tree in interval of -5 to 5.\
We can see that median is around 10, while 75% of data are from 0-20. And data more than 35 can be considered to be outliers.\
We can also see each outliers, and some special oueliers, like one of 450, one of 430 and so on, which is very big compared to other tree.\
Boxplot and histogram give us information about the distribution, but it is a general distribution for the whole data set. So maybe, now we want to know the distribution of each borough.


```{r}
library(ggridges)
ggplot(mydata, aes(x = tree_dbh, y = borough)) + geom_density_ridges()
```

Using a ridgeplot, we can compare the density distribution by borough very easily. Ridgeplot give us a direct understanding of different distribution, and easy to compare.\
We can get many information, for example, Queens has more trees with diameter from 25-50 than Manhattan, which have very few tree with diameter in that range. \
Also, we can found out a single-peak pattern in every borough, and a similar normal distributin of the pattern, which is useful in our further study if calculation will be involved.\

There are still many things we need to find out, for example, distribution for kinds of tree, relationship between borough and kinds and so on. We will contiune to explore the data in our final project. 
